# Summary

The following code was written as a part of the Complex Systems course at Tilburg University. The goal of the project is to analyse the complex coupling relationships between hands during piano play in different playing styles (namely 'normal', 'exaggerated' and 'still'). The 'expressive musical gestures' dataset reported in a paper by Sarasúa et al. was used to this end. The following code uses mdCRQA and transfer entropy to analyse the changing relationships between these playing styles. Grid search is used to optimize the parameters and surrogate testing is used to verify results are not due to chance.

Álvaro Sarasúa, Baptiste Caramiaux, Atau Tanaka, and Miguel Ortiz (2017). Datasets for the Analysis of Expressive Musical Gestures. In Proceedings of the 4th International Conference on Movement Computing, London, UK, June 2017 (MOCO'17).


# Setup
```{r setup, include=FALSE}
# R chunk
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
# R chunk
library(reticulate)
```


``` {r}
# R chunk
py_require("pandas")
py_require("multiSyncPy")
```

``` {python}
# Python chunk
import pandas as pd
import numpy as np
from glob import glob
import multiSyncPy as msp
from multiSyncPy import synchrony_metrics as sm
```

# Settings
``` {python}
# Python chunk

# You can replace any variable of the styles with * to include all variations
STYLES_OF_INTEREST = ["NOMETRO_NORMAL_NORMAL", "NOMETRO_NORMAL_EXAG", "NOMETRO_NORMAL_STILL"]
USE_VELOCITY = True # Set to 'False' to use raw coordinates
NORMALIZE = True # Also clips extreme values
LOWER_Q = 0.01 # Values below of these quantile bounds will be clipped to this quantile value
UPPER_Q = 0.99 # Values above of these quantile bounds will be clipped to this quantile value
PIANIST_NUM = "pianist01" # Selects which pianist is analysed. Must match folder name!

# Grid search hyperparameter tuning for mdCRQA
RR_RANGE = [1, 5] # Minimum, maximum values for RR for hyperparameters to be deemed appropriate.
MUST_MATCH_ALL = False # Determines whether all recordings for the selected style must match the same hyperparameters or whether they can differ.

DELAYS = [5,10,25] # Parameter value options for delay (d)
EMBEDS = [2,3,4,5,6] # Parameter value options for embedding dimension (m)
RADII = [0.1,0.2,0.3,0.4,0.5,0.65,0.8] # Parameter value options for radius (r)

# History length for transfer entropy
LXS = 5 # Number of past data points to consider for the history of the variable (max = 20)

# Random seed
SEED = 102026

``` 

# Load data
``` {python}
# Python chunk

labels = []

def load_data(styles = STYLES_OF_INTEREST, pianist=PIANIST_NUM):
  global labels
  data = []

  for style in STYLES_OF_INTEREST:
    file_names = glob(f"data/{pianist}/{style}_*_mocap.csv")
    labels += [file_name.split("/")[-1][:-10] for file_name in file_names]
    print(f"There are {len(file_names)} files for {style}.")
    for file_name in file_names:
      data.append(pd.read_csv(file_name, delimiter=";"))
      
  return [*data]
```

# Pre-processing
``` {python}
# Python chunk

def filter_data(data, q=USE_VELOCITY):
  df_raw = pd.DataFrame(data)
  df_clean = pd.DataFrame()

  # Select relevant columns
  if q: # calculate gradient if use_velocity is selected
    df_raw["LHAND_x"] = np.gradient(df_raw["LHAND_x"])
    df_raw["LHAND_y"] = np.gradient(df_raw["LHAND_y"])
    df_raw["LHAND_z"] = np.gradient(df_raw["LHAND_z"])
    
    df_raw["RHAND_x"] = np.gradient(df_raw["RHAND_x"])
    df_raw["RHAND_y"] = np.gradient(df_raw["RHAND_y"])
    df_raw["RHAND_z"] = np.gradient(df_raw["RHAND_z"])
    df_clean["LHAND_q"] = df_raw[["LHAND_x", "LHAND_y", "LHAND_z"]].values.tolist()
    df_clean["RHAND_q"] = df_raw[["RHAND_x", "RHAND_y", "RHAND_z"]].values.tolist()
  else: # else use 'raw' coordinates
    df_clean["LHAND"] = df_raw[["LHAND_x", "LHAND_y", "LHAND_z"]].values.tolist()
    df_clean["RHAND"] = df_raw[["RHAND_x", "RHAND_y", "RHAND_z"]].values.tolist()
    
  return df_clean

loaded_data = load_data()
filtered_data = [filter_data(l_data) for l_data in loaded_data]
print(labels)
```
There are 3 files for NOMETRO_NORMAL_NORMAL.
There are 3 files for NOMETRO_NORMAL_EXAG.
There are 3 files for NOMETRO_NORMAL_STILL.
['NOMETRO_NORMAL_NORMAL_2', 'NOMETRO_NORMAL_NORMAL_1', 'NOMETRO_NORMAL_NORMAL_3', 'NOMETRO_NORMAL_EXAG_3', 'NOMETRO_NORMAL_EXAG_1', 'NOMETRO_NORMAL_EXAG_2', 'NOMETRO_NORMAL_STILL_3', 'NOMETRO_NORMAL_STILL_2', 'NOMETRO_NORMAL_STILL_1']

# Normalization
``` {r}
# R chunk

# define normalization function
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
}

# This function handles min-max normalization as well as quantile capping as described in settings. If 'NORMALIZE' is set to false, this function is never ran and the original data is used.
normalize_data <- function(LHAND, RHAND) {
  LHAND <- as.data.frame(LHAND)
  RHAND <- as.data.frame(RHAND)
  
  # Quantile capping
  RHAND$x[RHAND$x < quantile(RHAND$x, py$LOWER_Q)] <- quantile(RHAND$x, py$LOWER_Q)
  RHAND$y[RHAND$y < quantile(RHAND$y, py$LOWER_Q)] <- quantile(RHAND$y, py$LOWER_Q)
  RHAND$z[RHAND$z < quantile(RHAND$z, py$LOWER_Q)] <- quantile(RHAND$z, py$LOWER_Q)
  RHAND$x[RHAND$x > quantile(RHAND$x, py$UPPER_Q)] <- quantile(RHAND$x, py$UPPER_Q)
  RHAND$y[RHAND$y > quantile(RHAND$y, py$UPPER_Q)] <- quantile(RHAND$y, py$UPPER_Q)
  RHAND$z[RHAND$z > quantile(RHAND$z, py$UPPER_Q)] <- quantile(RHAND$z, py$UPPER_Q)
  
  LHAND$x[LHAND$x < quantile(LHAND$x, py$LOWER_Q)] <- quantile(LHAND$x, py$LOWER_Q)
  LHAND$y[LHAND$y < quantile(LHAND$y, py$LOWER_Q)] <- quantile(LHAND$y, py$LOWER_Q)
  LHAND$z[LHAND$z < quantile(LHAND$z, py$LOWER_Q)] <- quantile(LHAND$z, py$LOWER_Q)
  LHAND$x[LHAND$x > quantile(LHAND$x, py$UPPER_Q)] <- quantile(LHAND$x, py$UPPER_Q)
  LHAND$y[LHAND$y > quantile(LHAND$y, py$UPPER_Q)] <- quantile(LHAND$y, py$UPPER_Q)
  LHAND$z[LHAND$z > quantile(LHAND$z, py$UPPER_Q)] <- quantile(LHAND$z, py$UPPER_Q)
  
  # Min-max normalization
  both_hands <- c(as.data.frame(LHAND), as.data.frame(RHAND))
  both_hands_norm <- as.data.frame(lapply(both_hands, min_max_norm))
  
  # Fixing variable names
  LHAND_new <- both_hands_norm[c('x', 'y', 'z')]
  colnames(LHAND_new) <- c("x","y","z")
  RHAND_new <- both_hands_norm[c('x.1', 'y.1', 'z.1')]
  colnames(RHAND_new) <- c("x","y","z")
    
  return(list(LHAND_new, RHAND_new))
}

normalized_data <- list()

for (data in py$filtered_data) {
  LHAND <- t(as.data.frame(data$LHAND))
  rownames(LHAND) <- NULL # To reset row names 
  colnames(LHAND) <- c("x","y","z")

  RHAND <- t(as.data.frame(data$RHAND))
  rownames(RHAND) <- NULL
  colnames(RHAND) <- c("x","y","z")
  
  if (py$NORMALIZE) {
    norm_data <- normalize_data(LHAND, RHAND)
    LHAND <- norm_data[1]
    RHAND <- norm_data[2]
  }
  
  LHAND <- as.data.frame(LHAND)
  RHAND <- as.data.frame(RHAND)
  
  normalized_data[[length(normalized_data)+1]] <- list(LHAND, RHAND)
}

# normalized_data is now a list of lists, each containing the left and right hand of a selected playstyle recording.
# To split them up into the desired playstyle groups again, you can verify the number of recordings each style
# has in the printed text under the 'Preprocessing' code block.
```


# Exploratory plot of data
``` {r}
# R chunk
library(plot3D)

scatter3D(LHAND$x, LHAND$y, LHAND$z)
scatter3D(RHAND$x, RHAND$y, RHAND$z)
```

# Optimize parameters
``` {r}
# R chunk
library(crqa)
library(pracma)

print(sprintf("Compatible parameters per style for %s<RR<%s based on parameter grid",py$RR_RANGE[[1]], py$RR_RANGE[[2]]))
print("")

# Provides parameters that fit the RR bounds in Settings given a parameter grid for delays, embedding dimensions and radii
optimizeParam_mdCRQA <- function (rec_LHAND, rec_RHAND, rec_delay, rec_embed, rec_radius) {
  param_options <- list()
  for (delay in rec_delay) {
    for (embed_dim in rec_embed) {
      for (radius in rec_radius) {
        result <- crqa(ts1=rec_LHAND, ts2=rec_RHAND, delay=delay, embed=embed_dim, radius=radius, method="mdcrqa", normalize=2)
        # Uncomment the line below to print all outputs, even if they fall outside of the preset bounds for RR
        #print(sprintf("d=%s;e=%s;r=%s -> RR=%s;DET=%s", delay, embed_dim, radius, round(result$RR, 3), round(result$DET, 3)))
        
        # Stores parameter combinations that fit RR range
        if (result$RR > py$RR_RANGE[[1]] && result$RR < py$RR_RANGE[[2]]) {
          param_options[[length(param_options)+1]] <- sprintf("d=%s;e=%s;r=%s", delay, embed_dim, radius)
          
          # Print intermediate results that match RR range.
          if (!py$MUST_MATCH_ALL) {
            print(sprintf("d=%s;e=%s;r=%s -> RR=%s;DET=%s", delay, embed_dim, radius, round(result$RR, 3), round(result$DET, 3)))
          }
        }
      }
    }
  }
  return(param_options)
}

# Simple function to help print parameter options that fit all recordings in a style
print_params <- function (style, param_options_list) {
    print(paste("Matches all recordings for",style))
    print(str(Reduce(intersect, param_options_list)))
    param_options_list <- list()
}

prev_style <- ""
param_options_list <- list()
for (i in 1:length(normalized_data)) {
  rec_label <- array(py$labels)[i] # e.g. NOMETRO_NORMAL_NORMAL_1
  rec_style <- substr(rec_label, 1, nchar(rec_label)-2) # e.g. NOMETRO_NORMAL_NORMAL
  rec_data <- normalized_data[i]
  
  # Print the parameters that work for all recordings in a specific style
  if (prev_style != rec_style && prev_style != "") {
    print_params(prev_style, param_options_list)
    param_options_list <- list()
  }
  if (!py$MUST_MATCH_ALL) {
    print(rec_label)
  }
  
  
  param_options <- optimizeParam_mdCRQA(rec_LHAND=rec_data[[1]][[1]], rec_RHAND=rec_data[[1]][[2]], rec_delay=py$DELAYS, rec_embed=py$EMBEDS, rec_radius=py$RADII)
  param_options_list[[length(param_options_list)+1]] <- param_options

  prev_style <- rec_style
}
print_params(prev_style)

```

[1] "Compatible parameters per style for 1<RR<5 based on parameter grid"
[1] ""
[1] "NOMETRO_NORMAL_NORMAL_2"
[1] "d=5;e=2;r=0.4 -> RR=2.115;DET=70.767"
[1] "d=5;e=2;r=0.5 -> RR=3.511;DET=74.941"
[1] "d=5;e=3;r=0.5 -> RR=1.621;DET=74.035"
[1] "d=5;e=3;r=0.65 -> RR=2.972;DET=79.381"
[1] "d=5;e=3;r=0.8 -> RR=4.701;DET=81.609"
[1] "d=5;e=4;r=0.65 -> RR=1.729;DET=77.933"
[1] "d=5;e=4;r=0.8 -> RR=2.789;DET=81.381"
[1] "d=5;e=5;r=0.65 -> RR=1.075;DET=76.934"
[1] "d=5;e=5;r=0.8 -> RR=1.811;DET=80.364"
[1] "d=5;e=6;r=0.8 -> RR=1.22;DET=79.895"
[1] "d=10;e=2;r=0.4 -> RR=1.875;DET=68.078"
[1] "d=10;e=2;r=0.5 -> RR=3.105;DET=73.265"
[1] "d=10;e=3;r=0.5 -> RR=1.307;DET=73.753"
[1] "d=10;e=3;r=0.65 -> RR=2.403;DET=77.935"
[1] "d=10;e=3;r=0.8 -> RR=3.819;DET=80.327"
[1] "d=10;e=4;r=0.65 -> RR=1.276;DET=78.155"
[1] "d=10;e=4;r=0.8 -> RR=2.085;DET=79.02"
[1] "d=10;e=5;r=0.8 -> RR=1.257;DET=80.455"
[1] "d=25;e=2;r=0.4 -> RR=1.461;DET=69.43"
[1] "d=25;e=2;r=0.5 -> RR=2.423;DET=74.093"
[1] "d=25;e=2;r=0.65 -> RR=4.235;DET=77.556"
[1] "d=25;e=3;r=0.65 -> RR=1.666;DET=78.708"
[1] "d=25;e=3;r=0.8 -> RR=2.606;DET=80.71"
[1] "d=25;e=4;r=0.8 -> RR=1.281;DET=79.403"
[1] "NOMETRO_NORMAL_NORMAL_1"
[1] "d=5;e=2;r=0.4 -> RR=1.689;DET=61.77"
[1] "d=5;e=2;r=0.5 -> RR=3.025;DET=67.716"
[1] "d=5;e=3;r=0.5 -> RR=1.153;DET=66.194"
[1] "d=5;e=3;r=0.65 -> RR=2.39;DET=69.217"
[1] "d=5;e=3;r=0.8 -> RR=4.206;DET=73.792"
[1] "d=5;e=4;r=0.65 -> RR=1.219;DET=70.628"
[1] "d=5;e=4;r=0.8 -> RR=2.158;DET=70.373"
[1] "d=5;e=5;r=0.8 -> RR=1.276;DET=71.968"
[1] "d=10;e=2;r=0.4 -> RR=1.5;DET=62.255"
[1] "d=10;e=2;r=0.5 -> RR=2.678;DET=68.056"
[1] "d=10;e=3;r=0.65 -> RR=1.957;DET=70.555"
[1] "d=10;e=3;r=0.8 -> RR=3.461;DET=73.519"
[1] "d=10;e=4;r=0.8 -> RR=1.666;DET=71.505"
[1] "d=25;e=2;r=0.4 -> RR=1.239;DET=61.032"
[1] "d=25;e=2;r=0.5 -> RR=2.208;DET=66.509"
[1] "d=25;e=2;r=0.65 -> RR=4.254;DET=72.628"
[1] "d=25;e=3;r=0.65 -> RR=1.513;DET=71.456"
[1] "d=25;e=3;r=0.8 -> RR=2.585;DET=75.235"
[1] "d=25;e=4;r=0.8 -> RR=1.134;DET=71.688"
[1] "NOMETRO_NORMAL_NORMAL_3"
[1] "Matches all recordings for NOMETRO_NORMAL_NORMAL"
 list()
NULL
[1] "NOMETRO_NORMAL_EXAG_3"
[1] "d=5;e=2;r=0.4 -> RR=2.061;DET=65.503"
[1] "d=5;e=2;r=0.5 -> RR=3.593;DET=69.181"
[1] "d=5;e=3;r=0.5 -> RR=1.516;DET=71.859"
[1] "d=5;e=3;r=0.65 -> RR=3.012;DET=72.619"
[1] "d=5;e=4;r=0.65 -> RR=1.661;DET=75.508"
[1] "d=5;e=4;r=0.8 -> RR=2.928;DET=76.654"
[1] "d=5;e=5;r=0.65 -> RR=1.083;DET=79.769"
[1] "d=5;e=5;r=0.8 -> RR=1.862;DET=78.596"
[1] "d=5;e=6;r=0.8 -> RR=1.312;DET=81.359"
[1] "d=10;e=2;r=0.4 -> RR=1.885;DET=65.714"
[1] "d=10;e=2;r=0.5 -> RR=3.28;DET=68.83"
[1] "d=10;e=3;r=0.5 -> RR=1.364;DET=72.625"
[1] "d=10;e=3;r=0.65 -> RR=2.71;DET=74.693"
[1] "d=10;e=3;r=0.8 -> RR=4.569;DET=76.857"
[1] "d=10;e=4;r=0.65 -> RR=1.468;DET=78.081"
[1] "d=10;e=4;r=0.8 -> RR=2.539;DET=78.126"
[1] "d=10;e=5;r=0.8 -> RR=1.588;DET=80.783"
[1] "d=10;e=6;r=0.8 -> RR=1.097;DET=83.216"
[1] "d=25;e=2;r=0.4 -> RR=1.618;DET=67.01"
[1] "d=25;e=2;r=0.5 -> RR=2.77;DET=70.358"
[1] "d=25;e=3;r=0.5 -> RR=1.066;DET=72.582"
[1] "d=25;e=3;r=0.65 -> RR=2.033;DET=76.094"
[1] "d=25;e=3;r=0.8 -> RR=3.407;DET=79.398"
[1] "d=25;e=4;r=0.65 -> RR=1.024;DET=77.849"
[1] "d=25;e=4;r=0.8 -> RR=1.738;DET=79.506"
[1] "NOMETRO_NORMAL_EXAG_1"
[1] "d=5;e=2;r=0.3 -> RR=1.134;DET=62.176"
[1] "d=5;e=2;r=0.4 -> RR=2.313;DET=67.22"
[1] "d=5;e=2;r=0.5 -> RR=3.86;DET=71.929"
[1] "d=5;e=3;r=0.5 -> RR=1.695;DET=71.172"
[1] "d=5;e=3;r=0.65 -> RR=3.21;DET=74.868"
[1] "d=5;e=4;r=0.65 -> RR=1.809;DET=73.207"
[1] "d=5;e=4;r=0.8 -> RR=3;DET=76.022"
[1] "d=5;e=5;r=0.65 -> RR=1.154;DET=77.179"
[1] "d=5;e=5;r=0.8 -> RR=1.926;DET=74.998"
[1] "d=5;e=6;r=0.8 -> RR=1.317;DET=77.474"
[1] "d=10;e=2;r=0.3 -> RR=1.025;DET=62.786"
[1] "d=10;e=2;r=0.4 -> RR=2.031;DET=67.551"
[1] "d=10;e=2;r=0.5 -> RR=3.375;DET=71.031"
[1] "d=10;e=3;r=0.5 -> RR=1.416;DET=71.054"
[1] "d=10;e=3;r=0.65 -> RR=2.656;DET=74.353"
[1] "d=10;e=3;r=0.8 -> RR=4.33;DET=78.197"
[1] "d=10;e=4;r=0.65 -> RR=1.408;DET=73.561"
[1] "d=10;e=4;r=0.8 -> RR=2.345;DET=76.094"
[1] "d=10;e=5;r=0.8 -> RR=1.421;DET=77.002"
[1] "d=25;e=2;r=0.4 -> RR=1.611;DET=67.674"
[1] "d=25;e=2;r=0.5 -> RR=2.68;DET=72.431"
[1] "d=25;e=2;r=0.65 -> RR=4.757;DET=77.088"
[1] "d=25;e=3;r=0.65 -> RR=1.825;DET=77.933"
[1] "d=25;e=3;r=0.8 -> RR=2.97;DET=78.395"
[1] "d=25;e=4;r=0.8 -> RR=1.451;DET=78.888"
[1] "NOMETRO_NORMAL_EXAG_2"
[1] "d=5;e=2;r=0.4 -> RR=1.057;DET=47.952"
[1] "d=5;e=2;r=0.5 -> RR=2.166;DET=56.415"
[1] "d=5;e=2;r=0.65 -> RR=4.642;DET=67.646"
[1] "d=5;e=3;r=0.65 -> RR=1.649;DET=60.818"
[1] "d=5;e=3;r=0.8 -> RR=3.291;DET=68.403"
[1] "d=5;e=4;r=0.8 -> RR=1.514;DET=65.04"
[1] "d=10;e=2;r=0.5 -> RR=1.904;DET=56.857"
[1] "d=10;e=2;r=0.65 -> RR=4.08;DET=67.641"
[1] "d=10;e=3;r=0.65 -> RR=1.288;DET=58.489"
[1] "d=10;e=3;r=0.8 -> RR=2.632;DET=66.586"
[1] "d=10;e=4;r=0.8 -> RR=1.063;DET=60.366"
Warning in max(t1) : no non-missing arguments to max; returning -Inf
[1] "d=25;e=2;r=0.5 -> RR=1.417;DET=55.845"
[1] "d=25;e=2;r=0.65 -> RR=3.133;DET=66.047"
[1] "d=25;e=3;r=0.8 -> RR=1.67;DET=64.626"
[1] "Matches all recordings for NOMETRO_NORMAL_EXAG"
List of 10
 $ : chr "d=5;e=2;r=0.4"
 $ : chr "d=5;e=2;r=0.5"
 $ : chr "d=5;e=3;r=0.65"
 $ : chr "d=5;e=4;r=0.8"
 $ : chr "d=10;e=2;r=0.5"
 $ : chr "d=10;e=3;r=0.65"
 $ : chr "d=10;e=3;r=0.8"
 $ : chr "d=10;e=4;r=0.8"
 $ : chr "d=25;e=2;r=0.5"
 $ : chr "d=25;e=3;r=0.8"
NULL
[1] "NOMETRO_NORMAL_STILL_3"
[1] "d=5;e=2;r=0.4 -> RR=1.208;DET=41.224"
[1] "d=5;e=2;r=0.5 -> RR=2.492;DET=56.174"
[1] "d=5;e=3;r=0.65 -> RR=1.931;DET=60.726"
[1] "d=5;e=3;r=0.8 -> RR=3.598;DET=69.677"
[1] "d=5;e=4;r=0.8 -> RR=1.726;DET=64.661"
[1] "d=10;e=2;r=0.4 -> RR=1.086;DET=39.592"
[1] "d=10;e=2;r=0.5 -> RR=2.252;DET=54.899"
[1] "d=10;e=2;r=0.65 -> RR=4.592;DET=67.223"
[1] "d=10;e=3;r=0.65 -> RR=1.57;DET=57.388"
[1] "d=10;e=3;r=0.8 -> RR=2.975;DET=66.806"
[1] "d=10;e=4;r=0.8 -> RR=1.256;DET=58.882"
[1] "d=25;e=2;r=0.5 -> RR=1.807;DET=52.732"
[1] "d=25;e=2;r=0.65 -> RR=3.792;DET=64.411"
[1] "d=25;e=3;r=0.65 -> RR=1.004;DET=50.367"
[1] "d=25;e=3;r=0.8 -> RR=2.068;DET=61.339"
[1] "NOMETRO_NORMAL_STILL_2"
[1] "d=5;e=2;r=0.5 -> RR=1.829;DET=54.843"
[1] "d=5;e=2;r=0.65 -> RR=3.793;DET=64.581"
[1] "d=5;e=3;r=0.65 -> RR=1.359;DET=56.977"
[1] "d=5;e=3;r=0.8 -> RR=2.584;DET=65.985"
[1] "d=5;e=4;r=0.8 -> RR=1.173;DET=59.936"
[1] "d=10;e=2;r=0.5 -> RR=1.671;DET=52.419"
[1] "d=10;e=2;r=0.65 -> RR=3.502;DET=63.448"
[1] "d=10;e=3;r=0.65 -> RR=1.128;DET=53.416"
[1] "d=10;e=3;r=0.8 -> RR=2.21;DET=62.881"
[1] "d=25;e=2;r=0.5 -> RR=1.277;DET=49.291"
[1] "d=25;e=2;r=0.65 -> RR=2.764;DET=61.223"
[1] "d=25;e=2;r=0.8 -> RR=4.904;DET=67.776"
[1] "d=25;e=3;r=0.8 -> RR=1.28;DET=55.327"
Warning in max(t1) : no non-missing arguments to max; returning -Inf
[1] "NOMETRO_NORMAL_STILL_1"
[1] "d=5;e=2;r=0.4 -> RR=1.373;DET=35.028"
[1] "d=5;e=2;r=0.5 -> RR=2.926;DET=50.466"
[1] "d=5;e=3;r=0.65 -> RR=2.149;DET=51.828"
[1] "d=5;e=3;r=0.8 -> RR=4.379;DET=64.437"
[1] "d=5;e=4;r=0.8 -> RR=1.933;DET=55.887"
[1] "d=10;e=2;r=0.4 -> RR=1.146;DET=30.935"
[1] "d=10;e=2;r=0.5 -> RR=2.512;DET=46.336"
[1] "d=10;e=3;r=0.65 -> RR=1.693;DET=46.675"
[1] "d=10;e=3;r=0.8 -> RR=3.593;DET=61.078"
[1] "d=10;e=4;r=0.8 -> RR=1.438;DET=49.444"
[1] "d=25;e=2;r=0.5 -> RR=2.05;DET=45.418"
[1] "d=25;e=2;r=0.65 -> RR=4.571;DET=62.062"
[1] "d=25;e=3;r=0.65 -> RR=1.169;DET=45.615"
[1] "d=25;e=3;r=0.8 -> RR=2.519;DET=60.468"
Warning in max(t1) : no non-missing arguments to max; returning -Inf
[1] "Matches all recordings for NOMETRO_NORMAL_STILL"
List of 10
 $ : chr "d=5;e=2;r=0.5"
 $ : chr "d=5;e=3;r=0.65"
 $ : chr "d=5;e=3;r=0.8"
 $ : chr "d=5;e=4;r=0.8"
 $ : chr "d=10;e=2;r=0.5"
 $ : chr "d=10;e=3;r=0.65"
 $ : chr "d=10;e=3;r=0.8"
 $ : chr "d=25;e=2;r=0.5"
 $ : chr "d=25;e=2;r=0.65"
 $ : chr "d=25;e=3;r=0.8"
NULL

# Results
## mdCRQA
```{r}
# R chunk

# NOTE: these values can actually be different from style to style! Therefore, feel free to run the code on several different hyperparameters and compare outputs of different styles on different hyperparameters.
# Change these based on the output of the previous code block. Ideally, the parameters 'fit' all recordings within a style
library(crqa)

DELAY <-      10
EMBED_DIM <-  3
RADIUS <-     0.8

# Print %RR and %DET for all recordings using selected parameters
print(sprintf("delay=%s; embed_dim=%s; radius=%s", DELAY, EMBED_DIM, RADIUS))
print("")
for (i in 1:length(normalized_data)) {
  rec_data <- normalized_data[i]

  result <- crqa(ts1=rec_data[[1]][[1]], ts2=rec_data[[1]][[2]], delay=DELAY, embed=EMBED_DIM, radius=RADIUS, method="mdcrqa", normalize=2)
  print(sprintf("%s - RR=%s; DET=%s", array(py$labels)[i], round(result$RR,3), round(result$DET,3)))
}
```
[1] "delay=10; embed_dim=3; radius=0.8"
[1] ""
[1] "NOMETRO_NORMAL_NORMAL_2 - RR=3.819; DET=80.327"
[1] "NOMETRO_NORMAL_NORMAL_1 - RR=3.461; DET=73.519"
[1] "NOMETRO_NORMAL_NORMAL_3 - RR=89.092; DET=99.999"
[1] "NOMETRO_NORMAL_EXAG_3 - RR=4.569; DET=76.857"
[1] "NOMETRO_NORMAL_EXAG_1 - RR=4.33; DET=78.197"
[1] "NOMETRO_NORMAL_EXAG_2 - RR=2.632; DET=66.586"
[1] "NOMETRO_NORMAL_STILL_3 - RR=2.975; DET=66.806"
[1] "NOMETRO_NORMAL_STILL_2 - RR=2.21; DET=62.881"
[1] "NOMETRO_NORMAL_STILL_1 - RR=3.593; DET=61.078"


## Transfer Entropy

### X coordinate
```{r}
library(RTransferEntropy)

#create a dataframe
all_res <- data.frame(i = numeric(), direction = character(), te = numeric(), ete = numeric(), se = numeric(), p=numeric(), lx =numeric())

#for all time series of interest
for (i in seq(normalized_data)) {
  rec_data <- normalized_data[i]

  #and for all history lengths up to LXS
  for (lx in seq(py$LXS)) {
    #compute transfer entropy
    te <- transfer_entropy(rec_data[[1]][[1]]$x,rec_data[[1]][[2]]$x, nboot = 100, lx=lx ,ly=lx, seed = py$SEED, quiet = TRUE)
    #add left->right and right->left TE values to the dataframe
    xy <- c(i, 'x->y', te$coef[1], te$coef[3], te$coef[5], te$coef[7], lx)
    yx <- c(i, 'y->x', te$coef[2], te$coef[4], te$coef[6], te$coef[8], lx)
    all_res <- rbind(all_res, xy)
    all_res <- rbind(all_res, yx)
    print(sprintf("for %s and history of: %s directionality index x-y: %s",array(py$labels)[i], lx, te$coef[1] - te$coef[2]))
  
  }
}
#make all appropriate columns numeric
colnames(all_res) = c('i','direction', 'te', 'ete', 'se', 'p', 'lx')
all_res$te <- as.numeric(as.character(all_res$te))
all_res$ete <- as.numeric(as.character(all_res$ete))
all_res$se <- as.numeric(as.character(all_res$se))
all_res$p <- as.numeric(as.character(all_res$p))
all_res$lx <- as.numeric(as.character(all_res$lx))
#Save only the significant values in a separate dataframe
sig_res <- subset(all_res, p < 0.05)
#display the significant results
print(sig_res)

#write significant results into a csv for easier interpretation
#write.csv(sig_res, "C:/Users/hajko/Desktop/Complex systems/project (new)/sig_xdata.csv", row.names=FALSE)
```
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 1 directionality index x-y: -0.00612502435196277"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 2 directionality index x-y: -0.00413808935624588"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 3 directionality index x-y: 0.000476685088071679"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 4 directionality index x-y: 0.00329003835323721"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 5 directionality index x-y: 0.00841517488563737"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 1 directionality index x-y: 0.0085754208621281"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 2 directionality index x-y: 0.00162043743019499"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 3 directionality index x-y: 0.00134783612452011"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 4 directionality index x-y: 0.00768558032977263"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 5 directionality index x-y: 0.00560298244183527"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 1 directionality index x-y: 0.00295512115844958"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 2 directionality index x-y: 0.00446870072343488"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 3 directionality index x-y: 0.0198568069554722"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 4 directionality index x-y: 0.00946298320363155"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 5 directionality index x-y: 0.0089871345766279"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 1 directionality index x-y: 0.00136791994744967"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 2 directionality index x-y: 0.0141891534585603"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 3 directionality index x-y: 0.020557710685927"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 4 directionality index x-y: 0.0256015413383466"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 5 directionality index x-y: 0.0253023965143232"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 1 directionality index x-y: 0.0115995595965608"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 2 directionality index x-y: 0.00847746302601586"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 3 directionality index x-y: 0.00944259258118482"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 4 directionality index x-y: 0.0151828424970183"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 5 directionality index x-y: 0.0234605059459319"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 1 directionality index x-y: 0.00694350587345545"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 2 directionality index x-y: 0.0109677131385426"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 3 directionality index x-y: 0.00690541145045934"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 4 directionality index x-y: 0.00240346381256114"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 5 directionality index x-y: 0.00789361040775152"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 1 directionality index x-y: 0.00751337643997649"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 2 directionality index x-y: 0.00245140100125357"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 3 directionality index x-y: 0.00118673733703131"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 4 directionality index x-y: 0.00777137544907669"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 5 directionality index x-y: 0.0135642372647959"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 1 directionality index x-y: 0.0089320036160402"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 2 directionality index x-y: 0.00504828056226321"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 3 directionality index x-y: 0.00565924666751155"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 4 directionality index x-y: 0.00619715660713965"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 5 directionality index x-y: 0.0114196663316373"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 1 directionality index x-y: -0.00903623391797406"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 2 directionality index x-y: -0.000131465506280543"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 3 directionality index x-y: -0.00265664469742841"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 4 directionality index x-y: 0.00152783757505574"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 5 directionality index x-y: 0.00199451976125693"
   i direction          te         ete          se    p lx
1  1      x->y 0.007750094 0.004213035 0.001503431 0.02  1
2  1      y->x 0.013875118 0.010716098 0.001202380 0.00  1
4  1      y->x 0.019895091 0.008036338 0.002982815 0.01  2
5  1      x->y 0.032396807 0.001992074 0.004915697 0.03  3
6  1      y->x 0.031920122 0.008203440 0.004063956 0.01  3
7  1      x->y 0.046778323 0.000000000 0.006487027 0.00  4
8  1      y->x 0.043488285 0.008773852 0.006828869 0.03  4
9  1      x->y 0.054900386 0.000000000 0.007739212 0.01  5
11 2      x->y 0.015404168 0.012236579 0.001199324 0.00  1
12 2      y->x 0.006828747 0.003571633 0.001221512 0.00  1
13 2      x->y 0.034465002 0.020021836 0.003090703 0.00  2
14 2      y->x 0.032844564 0.020252859 0.002922203 0.00  2
15 2      x->y 0.040868712 0.011461303 0.004752019 0.00  3
16 2      y->x 0.039520876 0.015932327 0.004970651 0.00  3
17 2      x->y 0.048296594 0.005236723 0.005217829 0.00  4
19 2      x->y 0.047990807 0.000000000 0.006049645 0.02  5
21 3      x->y 0.034985982 0.030783010 0.001299875 0.00  1
22 3      y->x 0.032030861 0.027848919 0.001477125 0.00  1
23 3      x->y 0.061051807 0.041983752 0.003627192 0.00  2
24 3      y->x 0.056583106 0.041605427 0.003426260 0.00  2
25 3      x->y 0.064841642 0.030645728 0.007045158 0.00  3
26 3      y->x 0.044984835 0.020305061 0.007403816 0.02  3
27 3      x->y 0.050326422 0.006316189 0.008829598 0.02  4
31 4      x->y 0.009861648 0.006499996 0.001455762 0.00  1
32 4      y->x 0.008493728 0.005533511 0.001158297 0.00  1
33 4      x->y 0.030497254 0.014476628 0.002942429 0.00  2
34 4      y->x 0.016308101 0.005214736 0.002758507 0.04  2
35 4      x->y 0.039880192 0.008688742 0.003868173 0.00  3
37 4      x->y 0.050301656 0.002235726 0.005716306 0.00  4
39 4      x->y 0.053598289 0.000000000 0.007067093 0.00  5
41 5      x->y 0.016744159 0.012661480 0.001424726 0.00  1
43 5      x->y 0.035387071 0.020236671 0.003228580 0.00  2
44 5      y->x 0.026909608 0.013881187 0.002911251 0.00  2
45 5      x->y 0.041748568 0.013356860 0.004854240 0.00  3
46 5      y->x 0.032305976 0.009567184 0.004780583 0.01  3
47 5      x->y 0.049713783 0.009100493 0.005781564 0.00  4
49 5      x->y 0.053559060 0.003851407 0.008008770 0.01  5
51 6      x->y 0.011847897 0.008289432 0.001424261 0.00  1
53 6      x->y 0.025871281 0.012235256 0.002504965 0.00  2
55 6      x->y 0.034121862 0.008123672 0.004716956 0.01  3
61 7      x->y 0.009302350 0.006591874 0.001155941 0.00  1
63 7      x->y 0.015632846 0.003029973 0.002393366 0.03  2
64 7      y->x 0.013181445 0.002862284 0.002044079 0.04  2
67 7      x->y 0.035977703 0.000000000 0.005661917 0.03  4
69 7      x->y 0.048128724 0.000000000 0.006299054 0.01  5
71 8      x->y 0.013083326 0.010250658 0.001274601 0.00  1
73 8      x->y 0.013431282 0.000000000 0.002133521 0.03  2
77 8      x->y 0.028288395 0.000000000 0.004436143 0.04  4
79 8      x->y 0.039495378 0.000000000 0.005637708 0.01  5
82 9      y->x 0.011200717 0.008375172 0.001245479 0.00  1
83 9      x->y 0.019332304 0.005067237 0.002456976 0.00  2
84 9      y->x 0.019463770 0.008313701 0.002323204 0.00  2
85 9      x->y 0.035187816 0.005089063 0.003927052 0.00  3
86 9      y->x 0.037844460 0.016005482 0.003393991 0.00  3
87 9      x->y 0.046100345 0.000000000 0.005459113 0.00  4
88 9      y->x 0.044572507 0.012711874 0.005506807 0.00  4
89 9      x->y 0.051698688 0.000000000 0.006939705 0.01  5
90 9      y->x 0.049704169 0.008382984 0.007371147 0.02  5

### Y coordinate
```{r}
# R chunk

library(RTransferEntropy)

all_res <- data.frame(i = numeric(), direction = character(), te = numeric(), ete = numeric(), se = numeric(), p=numeric(), lx =numeric())

for (i in seq(normalized_data)) {
  rec_data <- normalized_data[i]

  for (lx in seq(py$LXS)) {
    te <- transfer_entropy(rec_data[[1]][[1]]$y,rec_data[[1]][[2]]$y, nboot = 100, lx=lx ,ly=lx, seed = py$SEED, quiet = TRUE)
    xy <- c(i, 'x->y', te$coef[1], te$coef[3], te$coef[5], te$coef[7], lx)
    yx <- c(i, 'y->x', te$coef[2], te$coef[4], te$coef[6], te$coef[8], lx)
    all_res <- rbind(all_res, xy)
    all_res <- rbind(all_res, yx)
    print(sprintf("for %s and history of: %s directionality index x-y: %s",array(py$labels)[i], lx, te$coef[1] - te$coef[2]))
  
  }
}

colnames(all_res) = c('i','direction', 'te', 'ete', 'se', 'p', 'lx')
all_res$te <- as.numeric(as.character(all_res$te))
all_res$ete <- as.numeric(as.character(all_res$ete))
all_res$se <- as.numeric(as.character(all_res$se))
all_res$p <- as.numeric(as.character(all_res$p))
all_res$lx <- as.numeric(as.character(all_res$lx))
sig_res <- subset(all_res, p < 0.05)
print(sig_res)
```
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 1 directionality index x-y: 0.0169267124796554"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 2 directionality index x-y: -0.000137060883286577"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 3 directionality index x-y: 0.00753684617869713"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 4 directionality index x-y: 0.014525306153154"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 5 directionality index x-y: 0.0121981170380618"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 1 directionality index x-y: 0.0157837444475326"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 2 directionality index x-y: 0.00577526416530401"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 3 directionality index x-y: 0.000109672888474792"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 4 directionality index x-y: -0.00233325030154849"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 5 directionality index x-y: 0.00137769448799967"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 1 directionality index x-y: 0.0133411278766137"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 2 directionality index x-y: -0.00125301930504844"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 3 directionality index x-y: 0.00951446563909793"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 4 directionality index x-y: 0.00742081278334018"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 5 directionality index x-y: 0.00463151536579742"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 1 directionality index x-y: 0.00392536398546893"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 2 directionality index x-y: 0.00973350758684784"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 3 directionality index x-y: 0.00492079113684663"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 4 directionality index x-y: 0.00284846162474246"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 5 directionality index x-y: 0.0116172046389492"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 1 directionality index x-y: -0.0139875904230382"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 2 directionality index x-y: -0.0094328765605883"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 3 directionality index x-y: -0.0105924284588894"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 4 directionality index x-y: -0.00823494368486404"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 5 directionality index x-y: -0.0101693633461393"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 1 directionality index x-y: 0.00731908920492885"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 2 directionality index x-y: 0.00217581942103769"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 3 directionality index x-y: 0.000881907883486126"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 4 directionality index x-y: 0.00245853404925182"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 5 directionality index x-y: 0.00236516155803815"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 1 directionality index x-y: 0.0119288200459588"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 2 directionality index x-y: 0.00952351361297846"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 3 directionality index x-y: 0.0109623013033658"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 4 directionality index x-y: 0.0126724311200288"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 5 directionality index x-y: 0.0113538591373099"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 1 directionality index x-y: 0.00160537076722527"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 2 directionality index x-y: 0.00874145549177281"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 3 directionality index x-y: 0.010138842532014"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 4 directionality index x-y: 0.0145031916025306"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 5 directionality index x-y: 0.0149859183939609"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 1 directionality index x-y: -0.000631652434263233"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 2 directionality index x-y: 0.0022769003433068"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 3 directionality index x-y: 0.00682031121647124"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 4 directionality index x-y: 0.0117089912635793"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 5 directionality index x-y: 0.0165749541949963"
   i direction          te          ete           se    p lx
1  1      x->y 0.028311213 0.0248335455 0.0013204792 0.00  1
2  1      y->x 0.011384501 0.0078190831 0.0013835195 0.00  1
3  1      x->y 0.031074517 0.0179639895 0.0026437380 0.00  2
4  1      y->x 0.031211578 0.0169014112 0.0032781283 0.00  2
5  1      x->y 0.036365285 0.0111872749 0.0046527984 0.00  3
7  1      x->y 0.045031802 0.0067099165 0.0061325924 0.01  4
11 2      x->y 0.022025294 0.0185280818 0.0013988546 0.00  1
12 2      y->x 0.006241550 0.0025953275 0.0013588025 0.03  1
13 2      x->y 0.034226759 0.0207197870 0.0037409895 0.00  2
14 2      y->x 0.028451495 0.0123754699 0.0036588137 0.00  2
15 2      x->y 0.042875433 0.0176877955 0.0044708353 0.00  3
16 2      y->x 0.042765760 0.0077936562 0.0058835474 0.00  3
18 2      y->x 0.046660877 0.0000000000 0.0069684282 0.03  4
21 3      x->y 0.047845257 0.0436724719 0.0013966167 0.00  1
22 3      y->x 0.034504129 0.0299631170 0.0012966238 0.00  1
23 3      x->y 0.067448446 0.0504123746 0.0041565705 0.00  2
24 3      y->x 0.068701465 0.0520581316 0.0041958263 0.00  2
25 3      x->y 0.051813544 0.0216300260 0.0066288505 0.00  3
26 3      y->x 0.042299078 0.0161926977 0.0069224775 0.03  3
31 4      x->y 0.026966930 0.0229996300 0.0014040388 0.00  1
32 4      y->x 0.023041566 0.0193427086 0.0014702930 0.00  1
33 4      x->y 0.053106930 0.0332983688 0.0042899187 0.00  2
34 4      y->x 0.043373422 0.0254193381 0.0036974142 0.00  2
35 4      x->y 0.056066667 0.0181859816 0.0058426406 0.00  3
36 4      y->x 0.051145876 0.0157635077 0.0058553479 0.00  3
41 5      x->y 0.011151040 0.0069493990 0.0015832685 0.00  1
42 5      y->x 0.025138631 0.0212374255 0.0013785638 0.00  1
43 5      x->y 0.038417935 0.0206903791 0.0040586647 0.00  2
44 5      y->x 0.047850811 0.0293242209 0.0043666472 0.00  2
45 5      x->y 0.059115782 0.0233068078 0.0063815690 0.00  3
46 5      y->x 0.069708210 0.0340831122 0.0058347624 0.00  3
48 5      y->x 0.061473506 0.0122585731 0.0077946532 0.00  4
51 6      x->y 0.014907202 0.0113793425 0.0014320392 0.00  1
52 6      y->x 0.007588113 0.0038218042 0.0017924000 0.03  1
53 6      x->y 0.030947823 0.0156720017 0.0032547054 0.00  2
54 6      y->x 0.028772003 0.0122984360 0.0034339753 0.00  2
55 6      x->y 0.040089985 0.0079820816 0.0051346418 0.01  3
56 6      y->x 0.039208077 0.0050276213 0.0055538221 0.02  3
57 6      x->y 0.048856234 0.0018099524 0.0065222212 0.02  4
61 7      x->y 0.019683008 0.0170784301 0.0010522343 0.00  1
62 7      y->x 0.007754188 0.0046967285 0.0010858351 0.00  1
63 7      x->y 0.025730158 0.0140933172 0.0021576271 0.00  2
64 7      y->x 0.016206644 0.0063995242 0.0024134172 0.00  2
65 7      x->y 0.033845676 0.0110975295 0.0035198415 0.00  3
66 7      y->x 0.022883375 0.0026658991 0.0031758222 0.02  3
67 7      x->y 0.041816157 0.0060092599 0.0049832028 0.00  4
69 7      x->y 0.043649396 0.0000000000 0.0066600991 0.04  5
71 8      x->y 0.010337911 0.0075228007 0.0009678770 0.00  1
72 8      y->x 0.008732541 0.0058744370 0.0009618073 0.00  1
73 8      x->y 0.024085371 0.0127378511 0.0024984647 0.00  2
74 8      y->x 0.015343915 0.0044785868 0.0025863752 0.04  2
75 8      x->y 0.029732213 0.0079162908 0.0035497929 0.00  3
76 8      y->x 0.019593370 0.0004851213 0.0029421311 0.03  3
77 8      x->y 0.036657831 0.0035550299 0.0043494915 0.00  4
79 8      x->y 0.041089872 0.0000000000 0.0055391029 0.01  5
81 9      x->y 0.013302799 0.0101913230 0.0014244122 0.00  1
82 9      y->x 0.013934451 0.0106541638 0.0012352767 0.00  1
83 9      x->y 0.020950700 0.0084308853 0.0026429318 0.00  2
84 9      y->x 0.018673800 0.0074836668 0.0026596667 0.01  2
85 9      x->y 0.030515643 0.0070077336 0.0042666398 0.00  3
86 9      y->x 0.023695332 0.0037563094 0.0034498425 0.02  3
87 9      x->y 0.037885441 0.0013953429 0.0050879031 0.01  4
89 9      x->y 0.046344513 0.0000000000 0.0063348491 0.03  5


### Z coordinate
```{r}
# R chunk

library(RTransferEntropy)

all_res <- data.frame(i = numeric(), direction = character(), te = numeric(), ete = numeric(), se = numeric(), p=numeric(), lx =numeric())

for (i in seq(normalized_data)) {
  rec_data <- normalized_data[i]

  for (lx in seq(py$LXS)) {
    te <- transfer_entropy(rec_data[[1]][[1]]$z,rec_data[[1]][[2]]$z, nboot = 100, lx=lx ,ly=lx, seed = py$SEED, quiet = TRUE)
    xy <- c(i, 'x->y', te$coef[1], te$coef[3], te$coef[5], te$coef[7], lx)
    yx <- c(i, 'y->x', te$coef[2], te$coef[4], te$coef[6], te$coef[8], lx)
    all_res <- rbind(all_res, xy)
    all_res <- rbind(all_res, yx)
    print(sprintf("for %s and history of: %s directionality index x-y: %s",array(py$labels)[i], lx, te$coef[1] - te$coef[2]))
  
  }
}

colnames(all_res) = c('i','direction', 'te', 'ete', 'se', 'p', 'lx')
all_res$te <- as.numeric(as.character(all_res$te))
all_res$ete <- as.numeric(as.character(all_res$ete))
all_res$se <- as.numeric(as.character(all_res$se))
all_res$p <- as.numeric(as.character(all_res$p))
all_res$lx <- as.numeric(as.character(all_res$lx))
sig_res <- subset(all_res, p < 0.05)
print(sig_res)
```
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 1 directionality index x-y: 0.00682563194268446"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 2 directionality index x-y: 0.0057794883116728"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 3 directionality index x-y: 0.00299309068300098"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 4 directionality index x-y: 0.00453249897559799"
[1] "for NOMETRO_NORMAL_NORMAL_2 and history of: 5 directionality index x-y: 0.00183197143251411"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 1 directionality index x-y: -0.000388942092794851"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 2 directionality index x-y: 0.0101604532776306"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 3 directionality index x-y: 0.00948493271517642"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 4 directionality index x-y: 0.0124776021576484"
[1] "for NOMETRO_NORMAL_NORMAL_1 and history of: 5 directionality index x-y: 0.0180673407020213"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 1 directionality index x-y: 0.00238150523530792"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 2 directionality index x-y: 0.0116464829288024"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 3 directionality index x-y: 0.00442872140072128"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 4 directionality index x-y: 0.00183070317045932"
[1] "for NOMETRO_NORMAL_NORMAL_3 and history of: 5 directionality index x-y: 0.0063574589494951"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 1 directionality index x-y: -0.000575943317992263"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 2 directionality index x-y: 0.000901333749020926"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 3 directionality index x-y: -0.00139915418222463"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 4 directionality index x-y: 0.00175167811885667"
[1] "for NOMETRO_NORMAL_EXAG_3 and history of: 5 directionality index x-y: -0.00147802786784441"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 1 directionality index x-y: 0.00291302213058123"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 2 directionality index x-y: 0.000782134312011035"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 3 directionality index x-y: -0.000952328116068454"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 4 directionality index x-y: 0.00107252347002516"
[1] "for NOMETRO_NORMAL_EXAG_1 and history of: 5 directionality index x-y: 0.00177286722480842"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 1 directionality index x-y: 0.00516056607496697"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 2 directionality index x-y: 0.00402893092732281"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 3 directionality index x-y: -0.00395420868716098"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 4 directionality index x-y: 0.000865748606527857"
[1] "for NOMETRO_NORMAL_EXAG_2 and history of: 5 directionality index x-y: 0.00114020205115564"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 1 directionality index x-y: -0.00564834989773579"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 2 directionality index x-y: -0.0037907904073555"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 3 directionality index x-y: -0.00167221344686461"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 4 directionality index x-y: 0.00313836876876143"
[1] "for NOMETRO_NORMAL_STILL_3 and history of: 5 directionality index x-y: -0.00127836532140352"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 1 directionality index x-y: 0.00201077590252865"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 2 directionality index x-y: 0.00148732559896522"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 3 directionality index x-y: 0.000983178574652691"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 4 directionality index x-y: 0.00637132757645897"
[1] "for NOMETRO_NORMAL_STILL_2 and history of: 5 directionality index x-y: 0.00891781406024687"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 1 directionality index x-y: -0.000359958411367451"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 2 directionality index x-y: -0.00161191999662256"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 3 directionality index x-y: -0.00855150276747844"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 4 directionality index x-y: -0.0070182816618817"
[1] "for NOMETRO_NORMAL_STILL_1 and history of: 5 directionality index x-y: -0.00426659559852002"
   i direction          te         ete          se    p lx
1  1      x->y 0.011784301 0.008391519 0.001349718 0.00  1
3  1      x->y 0.020717462 0.007058343 0.002668283 0.00  2
11 2      x->y 0.011267330 0.007730137 0.001413193 0.00  1
12 2      y->x 0.011656272 0.008027173 0.001510757 0.00  1
13 2      x->y 0.031869364 0.018964533 0.003173783 0.00  2
14 2      y->x 0.021708911 0.008382324 0.003423247 0.03  2
15 2      x->y 0.030832412 0.004587261 0.004435812 0.03  3
21 3      x->y 0.034707448 0.030237665 0.001444256 0.00  1
22 3      y->x 0.032325942 0.027891188 0.001733129 0.00  1
23 3      x->y 0.063456168 0.046732607 0.004528685 0.00  2
24 3      y->x 0.051809685 0.035425460 0.003877914 0.00  2
25 3      x->y 0.056530656 0.028083940 0.006911950 0.00  3
26 3      y->x 0.052101934 0.021904897 0.006811344 0.00  3
31 4      x->y 0.008611074 0.005800626 0.001139109 0.00  1
32 4      y->x 0.009187018 0.006398135 0.001098306 0.00  1
41 5      x->y 0.005711649 0.002920753 0.001118245 0.01  1
43 5      x->y 0.012315189 0.002678155 0.002304053 0.02  2
44 5      y->x 0.011533054 0.001924067 0.002120291 0.01  2
51 6      x->y 0.008381095 0.004987775 0.001468675 0.00  1
62 7      y->x 0.009569664 0.007126413 0.001319996 0.00  1
64 7      y->x 0.018378719 0.005971971 0.002431869 0.02  2
71 8      x->y 0.009309987 0.006739873 0.001191631 0.00  1
72 8      y->x 0.007299211 0.004448757 0.001307996 0.00  1
73 8      x->y 0.018434146 0.007735453 0.002006333 0.00  2
75 8      x->y 0.031882370 0.009156430 0.003906493 0.00  3
76 8      y->x 0.030899191 0.006532353 0.004482117 0.02  3
77 8      x->y 0.046784246 0.010352676 0.006127759 0.00  4
86 9      y->x 0.035332679 0.005234780 0.005342065 0.01  3
87 9      x->y 0.044352999 0.004063642 0.005914751 0.04  4
88 9      y->x 0.051371280 0.003188101 0.006653709 0.01  4


# Surrogate Testing
``` {python}
# Python chunk

np.random.seed(SEED)

def generate_surrogates(ts, window_size, n_surrogates=100):
  """
  this code is adapted from the shuffle_time_windows function in Hudson, D., Wiltshire, T. J., &  Atzmueller, M. (2023). multiSyncPy: A Python package for assessing multivariate coordination dynamics. Behavior Research Methods, 55(2), 932-962. https://doi.org/10.3758/s13428-022-01855-y.
  
  parameters:
  ts: time series should be of shape (nrows, 3)
  window size: how to choose the size?
  
  returns: an array of generated surrogate data (by default 100 surrogate time series)
  """
  surrogates_list = []
  surrogate_shape = ts.shape
  ts = np.array(ts).copy()
  
  ## this trimming part can be commented out if needed
  # trim the data so that the data length is divisible by window_size
  n_windows = surrogate_shape[0] // window_size
  trimmed_size = n_windows * window_size
  ts = ts[:trimmed_size]
  ##
  
  for n in range(n_surrogates):
    # split the time series into windows
    surrogate_data = ts.reshape(-1, window_size, 3).copy() # so that the individual windows can be shuffled, not each individual row
    # shuffle the windows
    np.random.shuffle(surrogate_data)
    surrogate_data = surrogate_data.reshape(-1, 3)
    # add to the list of generated data
    surrogates_list.append(surrogate_data)
    
  return np.array(surrogates_list)

```

``` {python}
# Python chunk

all_surrogate_data = []
normalized_data = r.normalized_data

# save all surrogate data in one place, for left and right hand with playing intention label
for i in range(len(normalized_data)):
    intention = labels[i][15]  # the intention character (N or E)
    lhand = np.array(normalized_data[i][0])
    rhand = np.array(normalized_data[i][1])
    
    # generate and append surrogates for each hand
    all_surrogate_data.append({
        "intention": intention,
        "surrogatesL": generate_surrogates(lhand, window_size=200),  # shape: (100, trimmed_size, 3)
        "surrogatesR": generate_surrogates(rhand, window_size=200)   
    })
```

``` {r}
# R chunk

# create a data frame with all %RR and %DET values for surrogates, and corresponding intention labels
df <- data.frame(intention = character(), surr_RR = numeric(), surr_DET = numeric())

surr_num <- 100  # change this if necessary (should match n_surrogates)

# go over all surrogates in all_surrogate_data
for (i in 1:length(py$all_surrogate_data)) {
    surrogate_data <- py$all_surrogate_data[[i]]
    intention <- surrogate_data$intention
    surr_L <- surrogate_data$surrogatesL  # (100, trimmed_size, 3)
    surr_R <- surrogate_data$surrogatesR
  
    # access each one of the 100 generated surrogate series (for each hand) and perform mdCRQA  
    for (surr in 1:surr_num) {
        tsL_surr <- as.data.frame(surr_L[surr,,])
        tsR_surr <- as.data.frame(surr_R[surr,,])
        colnames(tsL_surr) <- c("x", "y", "z")
        colnames(tsR_surr) <- c("x", "y", "z")
        
        # mdCRQA with the current left and right hand surrogate series
        result <- crqa(ts1=tsL_surr, ts2=tsR_surr, delay=DELAY, embed=EMBED_DIM,
                       radius=RADIUS, method="mdcrqa", normalize=2)
        # save the %RR and %DET with the corresponding intention info
        df <- rbind(df, data.frame(intention=intention, surr_RR=result$RR,
                                   surr_DET=result$DET))
    }
}

# split into two datasets: normal and exaggerated playing
norm <- subset(df, intention == "N")
exag <- subset(df, intention == "E")
stil <- subset(df, intention == "S")
```

``` {r}
# R chunk

library(gmodels)
# Calculate mean
mean_RR <- mean(norm$surr_RR)
mean_RR
mean_DET <- mean(norm$surr_DET)
mean_DET
  
# Calculate length
n_RR <- length(norm$surr_RR)
n_DET <- length(norm$surr_DET)

# Calculate standard deviation
sd_RR <- sd(norm$surr_RR)
sd_DET <- sd(norm$surr_DET)

# Calculate Standard Error
error_RR <- qt(.95 + (1 - .95)/2, df = n_RR - 1) * sd_RR/sqrt(n_RR)
error_DET <- qt(.95 + (1 - .95)/2, df = n_DET - 1) * sd_DET/sqrt(n_DET)

# CI
CI_RR <- c(lower = mean_RR - error_RR, mean = mean_RR, upper = mean_RR + error_RR)
print(CI_RR)
CI_DET <- c(lower = mean_DET - error_DET, mean = mean_DET, upper = mean_DET + error_DET)
print(CI_DET)
```
[1] 30.28569
[1] 77.44411
   lower     mean    upper 
25.71647 30.28569 34.85491 
   lower     mean    upper 
75.60292 77.44411 79.28529 

(From mdCRQA)
[1] "delay=10; embed_dim=3; radius=0.8"
[1] ""
[1] "NOMETRO_NORMAL_NORMAL_2 - RR=3.819; DET=80.327"
[1] "NOMETRO_NORMAL_NORMAL_1 - RR=3.461; DET=73.519"
[1] "NOMETRO_NORMAL_NORMAL_3 - RR=89.092; DET=99.999"



```{r}
# R chunk

library(gmodels)
# Calculate mean
mean_RR <- mean(exag$surr_RR)
mean_RR
mean_DET <- mean(exag$surr_DET)
mean_DET
  
# Calculate length
n_RR <- length(exag$surr_RR)
n_DET <- length(exag$surr_DET)

# Calculate standard deviation
sd_RR <- sd(exag$surr_RR)
sd_DET <- sd(exag$surr_DET)

# Calculate Standard Error
error_RR <- qt(.95 + (1 - .95)/2, df = n_RR - 1) * sd_RR/sqrt(n_RR)
error_DET <- qt(.95 + (1 - .95)/2, df = n_DET - 1) * sd_DET/sqrt(n_DET)

# CI
CI_RR <- c(lower = mean_RR - error_RR, mean = mean_RR, upper = mean_RR + error_RR)
print(CI_RR)
CI_DET <- c(lower = mean_DET - error_DET, mean = mean_DET, upper = mean_DET + error_DET)
print(CI_DET)

```
[1] 3.127457
[1] 69.74703
   lower     mean    upper 
3.022799 3.127457 3.232115 
   lower     mean    upper 
68.84601 69.74703 70.64804

(From mdCRQA)
[1] "delay=10; embed_dim=3; radius=0.8"
[1] ""
[1] "NOMETRO_NORMAL_EXAG_3 - RR=4.569; DET=76.857"
[1] "NOMETRO_NORMAL_EXAG_1 - RR=4.33; DET=78.197"
[1] "NOMETRO_NORMAL_EXAG_2 - RR=2.632; DET=66.586"


```{r}
# R chunk

library(gmodels)
# Calculate mean
mean_RR <- mean(stil$surr_RR)
mean_RR
mean_DET <- mean(stil$surr_DET)
mean_DET
  
# Calculate length
n_RR <- length(stil$surr_RR)
n_DET <- length(stil$surr_DET)

# Calculate standard deviation
sd_RR <- sd(stil$surr_RR)
sd_DET <- sd(stil$surr_DET)

# Calculate Standard Error
error_RR <- qt(.95 + (1 - .95)/2, df = n_RR - 1) * sd_RR/sqrt(n_RR)
error_DET <- qt(.95 + (1 - .95)/2, df = n_DET - 1) * sd_DET/sqrt(n_DET)

# CI
CI_RR <- c(lower = mean_RR - error_RR, mean = mean_RR, upper = mean_RR + error_RR)
print(CI_RR)
CI_DET <- c(lower = mean_DET - error_DET, mean = mean_DET, upper = mean_DET + error_DET)
print(CI_DET)

```
[1] 2.228464
[1] 61.02928
   lower     mean    upper 
2.201228 2.228464 2.255700 
   lower     mean    upper 
60.86182 61.02928 61.19675 

(From mdCRQA)
[1] "delay=10; embed_dim=3; radius=0.8"
[1] ""
[1] "NOMETRO_NORMAL_STILL_3 - RR=2.975; DET=66.806"
[1] "NOMETRO_NORMAL_STILL_2 - RR=2.21; DET=62.881"
[1] "NOMETRO_NORMAL_STILL_1 - RR=3.593; DET=61.078"

